Characteristic polynomials of p-adic matrices

### 2017-02-02 charpoly/charpoly.tex

The second ingredient we need is the reciprocal polynomial.
We extend its definition to matrices of polynomials.
\begin{deftn}
Let $P \in K[X]$ be of degree $d \in \mathbb{N}_{>0}.$ 
We define its reciprocal polynomial as $P^{rec}=X^d P \left( 1/X \right).$
Let $A \in M_n(K[X]).$ We denote by $A^{rec}$ the matrix
whose coefficients are the reciprocal polynomials of the
coefficients of $A.$ 
\end{deftn}
We then have the following result :
\begin{lem}
Let $M \in M_n(K).$ Then:
\begin{eqnarray*}
com(1-XM)^{rec}=&com(X-M), \\
(\chi_M \times 1)^{rec}=&(1-XM) com(1-XM).\\
\end{eqnarray*}
\end{lem}

This lemma allows us to compute $com(1-XM)$ instead of $com(X-M).$
This has a remarkable advantage: the pivots during the computation of
the SNF of $com(1-XM)$ are units of $O_K[[X]],$ and are known
in advance to be on the diagonal. This leads to a very smooth
precision and complexity behaviour.

@@

\subsection{Computation and optimal jagged precision}

%Thanks to the previous results, we can now compute 
%$\com (X-M) \mod (\chi(M))$ in average complexity $\softO (n^3),$
%using the following algorithm.

>>>>

The second ingredient we need is the reciprocal polynomial.
We extend its definition to matrices of polynomials.
\begin{deftn}
Let $P \in K[X]$ be of degree $d \leq n.$ 
We define its reciprocal polynomial as $P^{rec}=X^d P \left( 1/X \right).$
Let $A \in M_n(K[X]).$ We denote by $A^{rec}$ the matrix
whose coefficients are the reciprocal polynomials of the
coefficients of $A.$ 
\end{deftn}
We then have the following result :
\begin{lem}
Let $M \in M_n(K).$ Then:
\begin{eqnarray*}
com(1-XM)^{rec}=&com(X-M), \\
(\chi_M \times 1)^{rec}=&(1-XM) com(1-XM).\\
\end{eqnarray*}
\end{lem}

This lemma allows us to compute $com(1-XM)$ instead of $com(X-M).$
This has a remarkable advantage: the pivots during the computation of
the SNF of $com(1-XM)$ are units of $O_K[[X]],$ and are known
in advance to be on the diagonal. This leads to a very smooth
precision and complexity behaviour when the entry matrix lives in 
$M_n(O_K).$

@@

\subsection{The comatrix of $X{-}M$}
